### 2024-04-17 docs/SOSP 2024/exploring.tex
Minor edits to exploration section

% \begin{itemize}
%   \item The scheduling language described in Section \ref{sec:schedule} allows
%   for numerous possible schedules for a given model. 
%   Finding a schedule with good performance is a non-trivial task.
%   \item We identify a reasonable template schedule for GPUs which encompasses
%   all strategies published in prior work.
%   \item Describe the template schedule.
%   \item Even within the variants of this template schedule, there is a significant 
%   amount of variation in performance. Show the histogram. 
%   \item Extensively exploring all possible parameter values for the template schedule 
%   is infeasible due to the large number of parameters.
%   Cite some numbers on how long it takes to explore the schedule space.
%   \item List the observations we make and the heuristic we design based on these.
% \end{itemize}
The set of schedules that can be constructed using the scheduling language described in 
Section \ref{sec:schedule} is unbounded.
\TODO{kr: next few lines need more work}
Even if one were to search for a good schedule offline, before deploying the model on a specific hardware, additional tools are needed help search for a 
high-performance schedule. We anticipate that even for the most complex models, one would like to spend at most a few minutes doing this search.
We propose a set of heuristics to guide the search for a good schedule and meet this goal. We do so in two steps, by first defining a bounded search space and then pruning this search space further.
\subsection{Bounding the search space}
We bound the space using a few meta parameters that together define the primitives used to construct a schedule. We do so while making sure that the space (at-least) covers the known strategies published in prior work. Specifically, we use 4 numeric parameters and 4 boolean parameters listed in Table \ref{tab:schedparams}. 
%Note that even this template has a large number ($1728$) of schedules, and requires further pruning as we discuss later.
The first three numeric parameters assign a configurable number of rows to each thread block, to each thread, and determine how trees are distributed across a specified number of threads. These parameters together define the loop schedule primitives to use (including the arguments to pass) from Table~\ref{Tab:LoopModifiers} and the parallelization strategy. The next four parameters determine the caching strategy, unroll factor{\footnote{We choose to always unroll trees completely, while its possible to partially unroll loops we did not observe any performance benefits from doing so.}} and how many trees to traverse simultaneously \TODO{Should we say, in the kernel/within a single thread?}.  (the remaining primitives from Table~\ref{Tab:Optimizations}).  
The final parameter determines what reduction options to use (Table~\ref{Tab:ReductionOpts}).
%Table \ref{tab:schedparams} lists the parameter values we tried. 
% \begin{itemize}
%   \item \textbf{Number of rows per thread block (Integer):} The number of rows that are processed by each thread block.
%   \item \textbf{Number of rows per thread (Integer):} The number of rows processed by each thread.
%   \item \textbf{Number of tree threads (Integer):} The number of threads across which the trees are distributed.
%   \item \textbf{Cache rows (Boolean):} Whether the input rows are cached in shared memory.
%   \item \textbf{Cache trees (Boolean):} Whether the trees are cached in shared memory.
%   \item \textbf{Unroll walks (Boolean):} Whether the tree walks are unrolled.
%   \item \textbf{Tree walk interleave factor:} The number of tree walks that are interleaved.
%   \item \textbf{Shared memory reduction:} Whether the reduction across tree threads is done in shared memory.
% \end{itemize}

@@

Exhaustive exploration over this bounded space ($1728$ schedules) is still very expensive, it took anywhere between thirty minutes to a few hours
to explore the entire space for a given model. 
%Exploring the space for every model is infeasible.
%\TODO{comment on online vs offline exploration? Set a goal for the heuristic?}
\subsection{Pruning the search space}
%Exploring the schedule space extensively even for a reasonable set of parameter values is very expensive.

@@

  %trees across more threads even at larger batch sizes. This is because processing fewer rows at a time 
  %allows us to keep them in shared memory.
  % We empirically find that the threshold for when 
  % we should start partitioning the trees across more threads is when the number of features
  % is greater than 100. 
  \subsubsection*{Orthogonal parameters} We find that some parameters like reduction type are orthogonal to the rest. 
  We therefore break the exploration into two phases, first we pick the best schedules without reduction and then evaluate reduction options on the best schedules.
    %We find that when a model prefers schedules with shared reduction, the same schedules 
  %without shared reduction are among the best performing schedules without shared reduction.
  % We therefore are able to separate the evaluation of shared reduction by collecting the 
  % best schedules without shared reduction and only evaluating shared reduction on them. 
  Evaluating the top 3 schedules for reduction is sufficient in practice.

>>>>

% \begin{itemize}
%   \item The scheduling language described in Section \ref{sec:schedule} allows
%   for numerous possible schedules for a given model. 
%   Finding a schedule with good performance is a non-trivial task.
%   \item We identify a reasonable template schedule for GPUs which encompasses
%   all strategies published in prior work.
%   \item Describe the template schedule.
%   \item Even within the variants of this template schedule, there is a significant 
%   amount of variation in performance. Show the histogram. 
%   \item Extensively exploring all possible parameter values for the template schedule 
%   is infeasible due to the large number of parameters.
%   Cite some numbers on how long it takes to explore the schedule space.
%   \item List the observations we make and the heuristic we design based on these.
% \end{itemize}
The set of schedules that can be constructed using the scheduling language described in 
Section \ref{sec:schedule} is unbounded.
\TODO{kr: next few lines need more work}
Even if one were to search for a good schedule offline, before deploying the model 
on a specific hardware, additional tools are needed to help search for a 
high-performance schedule. We anticipate that even for the most complex models, 
% one would like to spend at most a few minutes doing this search.
spending more than a few minutes on this search is not acceptable.
We propose a set of heuristics to guide the search for a good schedule 
and meet this goal. These heuristics work by first defining a 
bounded search space and then pruning this search space further
based on the batch size and model characteristics.

\subsection{Bounding the search space}
We bound the space using a few meta-parameters that together define 
the primitives used to construct a schedule. We do so while making sure 
that the space (at-least) covers the known strategies published in prior work.
Specifically, we use 4 numeric parameters and 4 boolean parameters listed in Table \ref{tab:schedparams}. 
%Note that even this template has a large number ($1728$) of schedules, and requires further pruning as we discuss later.
The first three numeric parameters assign a configurable number of rows to each 
thread block, to each thread, and determine how trees are distributed across a
specified number of threads. These parameters together define the loop schedule 
primitives to use (including the arguments to pass) from Table~\ref{Tab:LoopModifiers} 
and the parallelization strategy. The next four parameters determine the caching strategy, 
unroll factor{\footnote{We choose to always unroll trees completely, while its possible 
to partially unroll loops we did not observe any performance benefits from doing so.}} 
and how many trees to traverse simultaneously \TODO{Should we say, in the kernel/within
a single thread?}.  (the remaining primitives from Table~\ref{Tab:Optimizations}).  
The final parameter determines what reduction options to use (Table~\ref{Tab:ReductionOpts}).
%Table \ref{tab:schedparams} lists the parameter values we tried. 
% \begin{itemize}
%   \item \textbf{Number of rows per thread block (Integer):} The number of rows that are processed by each thread block.
%   \item \textbf{Number of rows per thread (Integer):} The number of rows processed by each thread.
%   \item \textbf{Number of tree threads (Integer):} The number of threads across which the trees are distributed.
%   \item \textbf{Cache rows (Boolean):} Whether the input rows are cached in shared memory.
%   \item \textbf{Cache trees (Boolean):} Whether the trees are cached in shared memory.
%   \item \textbf{Unroll walks (Boolean):} Whether the tree walks are unrolled.
%   \item \textbf{Tree walk interleave factor:} The number of tree walks that are interleaved.
%   \item \textbf{Shared memory reduction:} Whether the reduction across tree threads is done in shared memory.
% \end{itemize}

@@

Exhaustive exploration over this bounded space ($5184$ schedules) is still very expensive, taking
anywhere between thirty minutes to a few hours to explore the entire space for a given model. 
%Exploring the space for every model is infeasible.
%\TODO{comment on online vs offline exploration? Set a goal for the heuristic?}

\subsection{Pruning the search space}
%Exploring the schedule space extensively even for a reasonable set of parameter values is very expensive.

@@

  %trees across more threads even at larger batch sizes. This is because processing fewer rows at a time 
  %allows us to keep them in shared memory.
  % We empirically find that the threshold for when 
  % we should start partitioning the trees across more threads is when the number of features
  % is greater than 100. 
  \subsubsection*{Orthogonal parameters} We find that some parameters like reduction type are orthogonal to the rest. 
  We therefore break the exploration into two phases, first we pick the best schedules without reduction
  optimizations and then evaluate reduction options on the best schedules.
  %We find that when a model prefers schedules with shared reduction, the same schedules 
  %without shared reduction are among the best performing schedules without shared reduction.
  % We therefore are able to separate the evaluation of shared reduction by collecting the 
  % best schedules without shared reduction and only evaluating shared reduction on them. 
  Evaluating the top 3 schedules for reduction is sufficient in practice.
